# Result and Analysis {#result-analysis}

In this chapter, the result of three parts from the methodology is shown. The first section explains the Docker images and the second section is about the benchmark result. The analysis part is included directly for those sections.

## Docker Images

In this section, the exploration's result of each combination between R implementations and platforms for the R installation and geospatial packages installation is shown. The result is described per each R implementations. The overview of the result can be seen in figure \@ref(tab:vanillaRDockerImagesTable) for Docker image without geospatial R package and \@ref(tab:geospatialRDockerImagesTable) for Docker image with geospatial R package.
 
In those two tables, there are four kinds of results. The first one is **Yes** which means that the Docker image is successfully created and it works as expected for this theses. The second result is **No** which means that the Docker image is not possible to create properly. More detail of the failure is explained below. The third one is **Stop**. This result means the Docker image is not created because its geospatial R Docker image will not work properly based on the other Docker image result. And the last one is **?** which means, it may be possible to create the Docker image but is not created because of the time limit.
 
First, in order to gather all of the system dependencies required by geospatial R packages, R package `sysreqs` is used to access the `sysreqsdb` database. The result for each platform is obtained using this method. These results then used for creating geospatial R Docker images. The full result is available online^[https://github.com/ismailsunni/dockeRs/tree/master/scripts/sysreqs].

All Docker images except the TERR one is published in Docker hub^[https://hub.docker.com/u/ismailsunni] while all Docker file is published in Github repository^[https://github.com/ismailsunni/dockeRs/]. In the Github repository, there are both successful and unsuccessful Docker file. It also provides some notes about the detail of the error and some command to build, run, log in to the Docker container shell, and completeness check of the geospatial R packages.

```{r loadLibraryResultAnalysis, included=FALSE, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library('kableExtra')
library('microbenchmark')
library('altRnative')
```

```{r vanillaRDockerImagesTable, echo=FALSE}
vanillaRDockerImages <- data.frame(
  gnur = c('Yes', 'Yes', 'Yes'),
  mro = c('Yes', 'Yes', 'Yes'),
  renjin = c('Yes', 'Stop', 'Stop'),
  fastr = c('Yes', 'Yes', 'Stop'),
  pqr = c('Yes', 'Stop', 'Stop'),
  terr = c('Yes', 'Stop', 'Stop'),
  row.names = c('Debian', 'Fedora', 'Arch Linux')
  )
vanillaRDockerImages %>%
  kable(
    col.names = c('GNU R', 'MRO', 'Renjin', 'FastR', 'pqR', 'TERR'), 
    caption = "Vanilla R Docker Images",
    align = 'c') %>%
  kable_styling(latex_options = "striped") %>%
  row_spec(0, bold = T) %>%
  column_spec(1, bold = T)
```

```{r geospatialRDockerImagesTable, echo=FALSE}
geospatialRDockerImages <- data.frame(
  gnur = c('Yes', 'Yes', 'Yes'),
  mro = c('Yes', 'Yes', '?'),
  renjin = c('No', 'No', 'No'),
  fastr = c('No', 'No', 'No'),
  pqr = c('No', 'No', 'No'),
  terr = c('No', 'No', 'No'),
  row.names = c('Debian', 'Fedora', 'Arch Linux')
  )
geospatialRDockerImages %>%
  kable(
    col.names = c('GNU R', 'MRO', 'Renjin', 'FastR', 'pqR', 'TERR'), 
    caption = "Geospatial R Docker Images",
    align = 'c') %>%
  kable_styling(latex_options = "striped") %>%
  row_spec(0, bold = T) %>%
  column_spec(1, bold = T)
```

### GNU R

#### Debian

There is a Docker image available for GNU R on Debian^[https://hub.docker.com/_/r-base], so there is no need to work on this. This Docker image is created by Rocker Project^[https://www.rocker-project.org/]. This project also provides a GNU R Docker image with geospatial package^[https://hub.docker.com/r/rocker/geospatial]. Unfortunately, this Docker image does not contain all geospatial R package but only for a subset of it. The selected packages are the one which is slow or tricky to install and the one that has a general-purpose. The geospatial R Docker image for this thesis is created by extending the Rocker Project geospatial R Docker image (adding the missing package). In the end, both the Docker image for GNU R on Debian and its geospatial version can be created.

#### Fedora

By using the Docker image based on Fedora 32, the Docker image for GNU R on Fedora is created. Based on this Docker image, GNU R on Fedora with geospatial is created. There was a problem with installing system dependencies from `sysreqsdb`. The problem is `proj-epsg`, `proj-nad`, and `v8-314-devel` are no longer available. This problem is solved by removing the first two packages and changing the third on to `v8`. In the end, both the Docker image for GNU R on Fedora and its geospatial version can be created.

#### Arch Linux

Based on the `archlinux:20200106` Docker image, the GNU R Docker image on Arch Linux is created. From this Docker image, GNU R geospatial Docker image is created. There is a problem with installing system dependencies since not all of the system dependencies are available in the Arch Linux package database. These system dependencies can be installed from the Arch User Repository (AUR)^[https://aur.archlinux.org/]. To install a package from AUR on Docker image, it needs to create a user without a password. In the end, both the Docker image for GNU R on Fedora and its geospatial version can be created.
 
There is an interesting finding on Geospatial GNU R on the Arch Linux Docker image. There is a different behavior of R code in this Docker image compared to the one on Fedora and Debian. The difference is about the CRS comparison. It is caused by different versions of GDAL. The Fedora and Debian Docker image have GDAL 2.x while this Arch Linux has GDAL 3.x. The issue has been reported to `sf` issue tracker^[https://github.com/r-spatial/sf/issues/1238].

### Microsoft R Open

#### Debian

There is a Docker image available fro MRO 3.5.3 on Ubuntu 18.04 [@nustNuestMrodocker2019]. From this Docker image, MRO 3.5.3 with geospatial R packages is created. There are three problems found. The first one is `libav-tools` is no longer available on this Ubuntu version. It's changed with `ffmpeg`. Another issue is `gfortran` which is needed for some R packages is not found on the base image, so it should be installed first. The third one is the `polyclip` package can not be installed from within MRO^[https://github.com/ismailsunni/dockeRs/issues/32]. It is solved by installing the `polyclip` package from the source. In the end, both the Docker image for MRO 3.5.3 on Ubuntu and its geospatial version can be created.

#### Fedora

There is no Docker image for MRO on Fedora. At first, it's created based on the latest version of Fedora (Fedora 32). It works properly until `sf` is tried to install. There is an error that mentioned `proj_api.h not found in standard or given locations`^[https://github.com/ismailsunni/dockeRs/issues/39]. In the same Docker image, GNU R is installed to check whether the issue is MRO specific or the Docker image specific. It turns out that it is possible to install `sf` on that Docker image for GNU R. In the end, it is solved by using an older version of Fedora (Fedora 30). One possible explanation is that Fedora 32 has `proj` version 6.2 while Fedora 30 has `proj` version 5.2. MRO 3.5.3 is based on GNU R 3.5.3 which has better support for older `proj` version. In the end, both the Docker image for MRO 3.5.3 on Fedora 30 and its geospatial version can be created.

#### Arch Linux

The is no official support for Arch Linux from MRO but there is an MRO installer from AUR^[https://aur.archlinux.org/packages/microsoft-r-open/]. By using this package, a Docker image for MRO 3.5.3 on Arch Linux with tag 20200106 is created. Unfortunately, the Docker image for MRO 3.5.3 with the geospatial R package is not possible to create using this base image. The reason is that in MRO installed the pinned version of `sf` (based on R 3.5.3) on the snapshot. It expects to have GDAL 2.x which still has `pcs.csv`. This file is removed in GDAL 3.0 which is the current version in the latest Arch Linux.
 
There is a possibility to downgrade the installed GDAL version to match with the requirement of `sf` in MRO 3.5.3 snapshot. Unfortunately, since the discovery of the MRO installer on Arch Linux is very late, the geospatial Docker creation is not done due to the time limit. In the end, the Docker image for MRO 3.5.3 on Arch Linux is created but its geospatial version is not created due to time limit.

### Renjin

#### Debian

Renjin Docker image for Debian is available [@nustNuestRenjindockerDockerfiles2019]. Unfortunately, Renjin doesn't support `sf` [@RenjinOrgSf] because Renjin does not support `units` package [@RenjinOrgUnits] which is needed by `sf`. `units` package is not installed even though `udunits2` is available because Renjin does not support packages that depend on external libraries, in this case, `udunits2`^[https://stackoverflow.com/questions/58996151/rccp-h-not-found-in-renjin-when-compiling-package/59017512]. In the end, the Docker image for Renjin on Debian is created but its geospatial version can not be created.
 
#### Fedora
 
Since it is not possible to install geospatial packages on Renjin on Debian, the same situation applies for Renjin on Fedora. Because of this reason, the Docker image for Renjin on Fedora is not created and marked as _stop_ in the final result.
 
#### Arch Linux
 
Since it is not possible to install geospatial packages on Renjin on Debian, the same situation applies for Renjin on Arch Linux. Because of this reason, the Docker image for Renjin on Arch Linux is not created and marked as _stop_ in the final result.

### FastR

#### Debian

There is a Docker image for FastR 3.4.0 on Debian[@nustNuestFastrdocker2019]. But there is a problem with installing the `sf` package (one of the core geospatial packages). The problem is related to `proj` not found although it's already installed ^[https://github.com/oracle/fastr/issues/116]. The next attempt is using the latest FastR release 3.6.0, by extending an example from ^[https://github.com/graalvm/examples/issues/6]. Unfortunately, the same problem still exists. To check whether this problem is FastR specific or not, GNU R is installed for both Docker images and it's possible to install `sf` on it. This means the problem is specific to FastR. In parallel, a Docker image for FastR on Fedora is tried to create, but it is not possible. A more detail explanation is shown in the Fedora part below. In the end, it is possible to create a Docker image for FastR on Debian but it is not possible to create its geospatial version on Debian.
 
#### Fedora
 
Oracle Linux 7 has the same package manager as Fedora, so Oracle Linux 7 is also qualified as the Fedora-based Docker image in this research. Using the official GraalVM Docker image [@OracleGraalvmceDocker] based on Oracle Linux 7 it is possible to install FastR 3.6.1. But geospatial packages like `sf` is not possible to install because the GDAL is too old on Oracle Linux 7. It is solved by building GDAL from source^[https://gist.github.com/simondobner/f859b2db15ad65090c3c316d3c224f45]. A similar problem (old package) also happens with `Pandoc` that is used for running the book knitting in SDSR. This is not mandatory though for geospatial workflow. 

Unfortunately, another problem occurs when running the R script from the SDSR book. It is found out that `sf` in FastR can not read a geopackage file. The problem has been reported to FastR repository[https://github.com/oracle/fastr/issues/128], but there are no comments from the maintainer or other people. From these results, it is concluded that it is not possible to install geospatial packages properly on FastR. In the end, it is possible to create a Docker image for FastR on Fedora but it is not possible to create its geospatial version on Fedora.
 
#### Arch Linux
 
Since it is not possible to install geospatial packages on FastR on Debian and Fedora, the same situation applies for FastR on Arch Linux. Because of this reason, the Docker image for FastR on Arch Linux is not created and marked as _stop_ in the final result.

### pqR

#### Debian

There is already a pqR Docker image for Debian[@nustNuestPqrdocker2019]. Unfortunately, the latest version of pqR is based on GNU R 2.15.0 [@m.nealPqRPrettyQuick2019]. This makes geospatial packages like `sf` not compatible with it since `sf` needs R version > 3.3.0.[@pebesmaSfSimpleFeatures2019]. Other examples of R packages that are no compatible are `stars`, `RQGIS`, and the latest of `sp`. `sf` is a mandatory package to run the geospatial workflow on the benchmark step. Without this package, it is not possible to run the benchmark. Therefore, the geospatial Docker image for pqR is not possible to create although it is possible to create the vanilla version.
 
#### Fedora

Since it is not possible to install geospatial packages on pqR on Debian, the same situation applies for pqR on Fedora. Because of this reason, the Docker image for pqR on Fedora is not created and marked as _stop_ in the final result.

#### Arch Linux

Since it is not possible to install geospatial packages on pqR on Debian, the same situation applies for pqR on Arch Linux. Because of this reason, the Docker image for pqR on Arch Linux is not created and marked as _stop_ in the final result.


### TIBCO Enterprise Runtime for R (TERR) {#docker-image-terr}

#### Debian

Although TERR is not free and open, TIBCO provides TERR for free with a Developer Evaluation license. This license has a limitation which prevents publishing the Docker image of TERR^[https://tap.tibco.com/storefront/product-view.ep?pID=15307]. With this limitation, this research only provides Dockerfile to build the Docker image that needs the installer that can be obtained from the TIBCO website. TERR is successfully installed for Debian and Ubuntu. Unfortunately, there is a problem when R geospatial packages are tried to be installed. If there is already a package installed, it is not possible to re-install the package. For example, package `vctrs` is already installed from another R packages installation, if we want to install `RSQLite` which depends on `vctrs` it will turn an error.
 
Another problem found is related to the `rJava` package. An error message suggests running `R CMD javareconf` for successful `rJava` installation. Unfortunately, this `javareconf` is not available in TERR.
 
From this problem, it is not possible to install geospatial packages that have the same R package dependency. In the end, it is possible to create a Docker image for TERR on Debian but it is not possible to create its geospatial Docker image.

#### Fedora

Since it is not possible to install geospatial packages on TERR on Debian, the same situation applies for TERR on Fedora. Because of this reason, the Docker image for TERR on Fedora is not created and marked as _stop_ in the final result.

#### Arch Linux

Since it is not possible to install geospatial packages on TERR on Debian, the same situation applies for TERR on Arch Linux. Because of this reason, the Docker image for TERR on Arch Linux is not created and marked as _stop_ in the final result.

## Benchmark {#benchmark}

This section is divided into three parts. The first one is a brief explanation of the benchmarking tool that is created and used in this research. The second part is the result and analysis of the baseline benchmark. And the last one is the result and analysis for the geospatial workflow benchmark.

### Benchmark Tool (`altRnative`)

A new R package name `altRnative` is created to provide functionalities to do benchmark for R script across different Docker image [@altRnative2020]. `altRnative` uses `microbenchmark` [@mersmannMicrobenchmarkAccurateTiming2019] for running the benchmark process and `stevedore` [@fitzjohnStevedoreDockerClient2019] for managing the Docker image and containers. This package is available in the GitHub repository, in the following link: `https://github.com/ismailsunni/altRnative`.
 
In the current version, `altRnative` only supports the successful Docker image created in this research as shown in table \@ref(tab:geospatialRDockerImagesTable). Users can choose which platform and R implementation that will be used in the benchmarking. The R script target for the benchmark can be provided to the `code` parameter. It can accept a raw string that represents R code or an `expression`. For running R code from a file, it can use the `source` method. An example of a running benchmark can be seen below.

```{R baselineBenchmarkCode, eval=FALSE}
baseline_benchmark_result = benchmarks_code(
  code = "1 + 1", 
  r_implementations = c('gnu-r', 'mro'), 
  platforms = c('debian', 'ubuntu', 'fedora', 'archlinux'),
  times = 10
  )
```

After the benchmark result is obtained, it can plot using `microbenchmark`'s default plot or box plot. Other than that, `altRnative` provides its own box plot function for most common usage. The sample plot can be seen in section \@ref(baseline-benchmark) and \@ref(geospatial-workflow-benchmark).
 
Another useful feature of `altRnative` in calculating the elapsed time ratio between the benchmark result using `normalize_benchmark_result` function. This function gives the ratio for each benchmark result compared to the mean of selected Docker image (implementation and platform.). From this result, it is easier to see the comparison between Docker images. From this result, it also can be plotted using the default box plot method, or directly using `normalize_benchmark_boxplot` from `altRnative` that calculating the ratio internally.

### Baseline Benchmark {#baseline-benchmark}

```{r load_data, included=FALSE, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
baseline_benchmark = readRDS(file = "./data/simple_benchmark_result.rds")  # 1 + 1
benchmark_result_per_chapter = readRDS(file = "./data/benchmark_result.rds")  # per chapter
sdsr_code_benchmark = readRDS(file = "./data/sdsr_code_benchmark_result.rds") # all code in one run
nano_time_factor = 10^9
options(digits=3)
```

The baseline benchmark is used to check the time spent for running the Docker container and the R inside it. To do it, a very simple R script (`1 + 1`) is run 10 times for each Docker image. The summary can be seen at table \@ref(tab:summaryTableBaseline). It's clear that there is not so much difference between Docker images for this baseline benchmark. The distribution of the baseline benchmark is shown in box plot figure \@ref(fig:baseline-benchmark). There are two outlier observations, but in general, the result is quite similar within the 0.1-second range for the mean.

```{r create_summary_table_baseline, included=FALSE, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
summaryTableBaseline <- baseline_benchmark %>% 
      group_by(expr) %>% 
      summarise(mintime = min(time) / nano_time_factor,
                meantime = mean(time) / nano_time_factor,
                mediantime = median(time) / nano_time_factor,
                maxtime = max(time) / nano_time_factor
                )
```

```{r summaryTableBaseline, echo=FALSE, results='hold'}
summaryTableBaseline %>%
  kable(
    "latex",
    col.names = c("Docker Image", "Min", "Mean", "Median", "Max"),
    caption = "Baseline Benchmark",
    booktabs = T) %>%
  kable_styling(
    latex_options = "striped"
    )
```

```{r baseline-benchmark, echo=FALSE, fig.align='center', fig.cap='Baseline Benchmark', fig.width = 10, out.width=400, fig.show='hold'}
benchmark_boxplot(baseline_benchmark, main = 'Baseline Benchmark')
```

### Geospatial Workflow Benchmark {#geospatial-workflow-benchmark}

For running geospatial workflow, the code from the SDSR book is extracted using an R script to avoid overhead time in knitting the book process. The result of this benchmark can be seen in table \@ref(tab:summaryTableSDSR) and its distribution in box plot figure \@ref(fig:sdsr-benchmark). In this result, there is no outlier for all Docker images. We can see also that GNU R has better compared (~27-28 seconds) to MRO (~33 seconds) regardless of the operating system. For a better view of the speed comparison, a normalized time is created by using the mean of time for GNU R on Debian as the baseline. The result is shown in table \@ref(tab:normalizeSummaryTableSDSR). It's clear that MRO needs 1.25x time compared to GNU R.

```{r create_summary_table_sdsr, included=FALSE, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
summaryTableSDSR <- sdsr_code_benchmark %>% 
      group_by(expr) %>% 
      summarise(mintime = min(time) / nano_time_factor,
                meantime = mean(time) / nano_time_factor,
                mediantime = median(time) / nano_time_factor,
                maxtime = max(time) / nano_time_factor
                )
```

```{r create_summary_normalize_table_sdsr, included=FALSE, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
normalizeSummaryTableSDSR <- sdsr_code_benchmark %>% 
      normalize_benchmark_result('gnu-r on debian') %>%
      group_by(expr) %>% 
      summarise(mintime = min(time),
                meantime = mean(time),
                mediantime = median(time),
                maxtime = max(time)
                )
```

```{r summaryTableSDSR, results='hold', echo=FALSE}
summaryTableSDSR %>%
  kable(
    "latex",
    col.names = c("Docker Image", "Min", "Mean", "Median", "Max"),
    caption = "SDSR Benchmark",
    booktabs = T) %>%
  kable_styling(
    latex_options = "striped"
    )
```

```{r normalizeSummaryTableSDSR, results='hold', echo=FALSE}
normalizeSummaryTableSDSR %>%
  kable(
    "latex",
    col.names = c("Docker Image", "Min", "Mean", "Median", "Max"),
    caption = "Normalized SDSR Benchmark",
    booktabs = T) %>%
  kable_styling(
    latex_options = "striped"
    )
```

```{r sdsr-benchmark, echo=FALSE, fig.align='center', fig.cap='Duration for All Code SDSR Benchmark', fig.width = 10, out.width=400, fig.show='hold'}
benchmark_boxplot(sdsr_code_benchmark, main = 'Duration for All Code SDSR Benchmark')
```

```{r ratio-sdsr-benchmark, echo=FALSE, fig.align='center', fig.cap='Duration Ratio for All SDSR Code', fig.width = 10, out.width=400, fig.show='hide'}
normalize_benchmark_boxplot(
    sdsr_code_benchmark, 
    baseline_image = 'gnu-r on debian', 
    main = 'Duration Ratio for All SDSR Code'
    )
```

It also can be check which chapter of the book that takes the most time to run in table \@ref(tab:longestChapter). Chapter `04-Raster-Cube.R` takes the most of the time compared to other chapters with 26.01 seconds. From here we can examine the benchmark result of chapter `04-Raster-Cube.R`.
 
Based on table \@ref(tab:Chapter04SummaryTable) the difference between the Docker images are not big. It is more clear in the normalized benchmark table \@ref(tab:Chapter04NormalizedTable) that MRO needs around at most 1.2x time compared to the GNU R one. There is also no outlier in the benchmark `04-Raster-Cube.R` as shown in box plot figure \@ref(fig:Chapter04BoxPlot). Full result for benchmark per chapter can be found in appendix \@ref(apendix-benchmark-per-chapter).
 

```{r find_longest_chapter, included=FALSE, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library('plyr')
df_chapters <- ldply (benchmark_result_per_chapter, data.frame)
colnames(df_chapters) <- c('chapter', 'dockerimage', 'time')
# Important to unload the package
detach(package:plyr, unload=TRUE)
timePerChapter <- df_chapters %>% 
  select(chapter, time) %>% 
  group_by(chapter) %>% 
  summarise(meantime = mean(time) / nano_time_factor)
```

```{r longestChapter, results='hold', echo=FALSE}
timePerChapter %>%
  kable(
    "latex",
    col.names = c("Chapter", "Mean"),
    caption = "Duration per Chapter for All Docker Images",
    booktabs = T) %>%
  kable_styling(
    latex_options = "striped",
    full_width = T
    )
```

```{r createChapter04SummaryTable, included=FALSE, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
chapter04SummaryTable <- benchmark_result_per_chapter[["04-Raster-Cube.R"]] %>% 
      group_by(expr) %>% 
      summarise(mintime = min(time) / nano_time_factor,
                meantime = mean(time) / nano_time_factor,
                mediantime = median(time) / nano_time_factor,
                maxtime = max(time) / nano_time_factor
                )
```

```{r appendixtablechapter04, results='hold', echo=FALSE}
chapter04SummaryTable %>%
  kable(
    "latex",
    col.names = c("Docker Image", "Min", "Mean", "Median", "Max"),
    caption = "Chapter 04 Benchmark",
    booktabs = T) %>%
  kable_styling(
    latex_options = "striped"
    )
```

```{r createChapter04NormalizedTable, included=FALSE, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
chapter04NormalizedTable <- benchmark_result_per_chapter[["04-Raster-Cube.R"]] %>% 
      normalize_benchmark_result('gnu-r on debian') %>%
      group_by(expr) %>% 
      summarise(mintime = min(time),
                meantime = mean(time),
                mediantime = median(time),
                maxtime = max(time)
                )
```

```{r Chapter04NormalizedTable, results='hold', echo=FALSE}
chapter04NormalizedTable %>%
  kable(
    "latex",
    col.names = c("Docker Image", "Min", "Mean", "Median", "Max"),
    caption = "Normalized Chapter 04 Benchmark",
    booktabs = T) %>%
  kable_styling(
    latex_options = "striped"
    )
```

```{r Chapter04BoxPlot, echo=FALSE, fig.align='center', fig.cap='Chapter 04 Benchmark', fig.width = 10, out.width=400, fig.show='hold'}
benchmark_boxplot(benchmark_result_per_chapter[["04-Raster-Cube.R"]], main = 'Chapter 04 Benchmark')
```

```{r NormalizedChapter04BoxPlot, echo=FALSE, fig.align='center', fig.cap='Normalized Chapter 04 Benchmark', fig.width = 10, out.width=400, fig.show='hide'}
normalize_benchmark_boxplot(
    benchmark_result_per_chapter[["04-Raster-Cube.R"]], 
    baseline_image = 'gnu-r on debian', 
    main = 'Normalized Chapter 04 Benchmark'
    )
```
